Hydras \& Co.

### 2022-04-18 doc/gaia-chapter.tex
add an index for gaia bridge

\chapter{Gaia and the hydra (draft)}
\index{coq}{Plug-ins!Gaia}
\label{gaia-chapter}

@@

\subsection{Hessenberg sum}

Natural sum  (a.k.a. Hessenberg's  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

@@

\subsection{Canonical sequences}
\label{sect:gcanon}
Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
\href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}}.
Module~\href{../theories/html/gaia_hydras.GCanon.html}%
{\texttt{gaia\_hydras.GCanon}}, imports definitions and results from that library.

Like for Hessenberg's sum, we define a function \texttt{canon}
\emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

@@

\subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}
\label{sect:gpath}

@@

\section{Rapidly growing arithmetic functions}
In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
The type of such families is \texttt{E0 -> nat -> nat} where
\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
In order to compare such functions, we define a few abstract properties in
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).

>>>>

\chapter{Gaia and the hydra (draft)}
\index{coq}{Plug-ins!Gaia}
\index{gaiabridge}{Introduction}
\label{gaia-chapter}

@@

\subsection{Hessenberg sum}
\index{gaiabridge}{Hessenberg sum}
Natural sum  (a.k.a. Hessenberg  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

@@

\subsection{Canonical sequences}
\label{sect:gcanon}
\index{gaiabridge}{Canonical sequences}

Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
\href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}}.
Module~\href{../theories/html/gaia_hydras.GCanon.html}%
{\texttt{gaia\_hydras.GCanon}}, imports definitions and results from that library.

Like for Hessenberg sum, we define a function \texttt{canon}
\emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

@@

\subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}
\label{sect:gpath}
\index{gaiabridge}{Accessibility and paths inside $\epsilon_0$}

@@

\section{Rapidly growing arithmetic functions}
\index{gaiabridge}{Rapidly growing functions}

In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
The type of such families is \texttt{E0 -> nat -> nat} where
\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
In order to compare such functions, we define a few abstract properties in
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).

### 2022-04-18 doc/hydras.tex
add an index for gaia bridge

\title{Hydras \& Co.}
%\pagestyle{fancy}
%\fancyhead{Hydras, Ordinals, \& Co.}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\begin{titlepage}
   \begin{center}
       \vspace*{1cm}

@@

Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp) (please look at Sect.~\vref{sect:gaia-first-intro} and the paragraphs signalled with {\gaiasign}).

@@

\printindex{coq}{Coq, plug-ins and standard library}
\printindex{maths}{Mathematical notions and algorithmics}
%\printindex{hydras}{Library hydras: Ordinals and hydra battles}
%\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
%\printindex{additions}{Library additions: Addition chains}

>>>>

\title{Hydras \& Co.}
%\pagestyle{fancy}
%\fancyhead{Hydras, Ordinals, \& Co.}
\makeindex{gaiabridge}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\begin{titlepage}
   \begin{center}
       \vspace*{1cm}

@@

\index{gaiabridge}{Introduction}
Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp) (please look at Sect.~\vref{sect:gaia-first-intro} and the paragraphs signalled with {\gaiasign}).

@@

\printindex{gaiabridge}{Links to \gaia Library}
\label{gaia-index}
\printindex{coq}{Coq, plug-ins and standard library}
\printindex{maths}{Mathematical notions and algorithmics}
%\printindex{hydras}{Library hydras: Ordinals and hydra battles}
%\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
%\printindex{additions}{Library additions: Addition chains}

### 2022-04-18 doc/part-hydras.tex
additions to gaia chapter; remove several proof-based definitions
update gaia-chapter
add an index for gaia bridge

 \subsection*{Compatibility with \gaia (in progress)}
 \label{sect:gaia-first-intro}
The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
aims  to formalize mathematics in \coq  in
the style of Nicolas Bourbaki. It contains many definitions and results about ordinal numbers. In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa. Remarks about compatibility with \gaia are signalled with the picture \gaiasign.

@@

\begin{itemize}
\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}. This representation uses several (hopefully harmless) axioms. We use it as a reference for proving the correctness of ordinal notations.
  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}. Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.

@@

\section{Ordinal Notations and \gaia's ordinals}
\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).

@@

\subsection{About Gaia}
\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.

@@

\input{movies/snippets/T1/finiteOrds}

\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.

@@

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.

@@

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).

@@

\paragraph*{\gaiasign}
In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).

@@

\paragraph*{\gaiasign}
In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).

@@

\paragraph*{\gaiasign}
In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:

@@

\paragraph*{\gaiasign}
 The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.

@@

\paragraph*{\gaiasign}
In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.

@@

@@

    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
    addition. It is used as an auxiliary operation  for defining variants
    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.

@@

    \index{hydras}{Projects}
    \begin{project}
    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
    (\texttt{coq-equations}, \texttt{Function}, etc.).
    \end{project}
    \subsection{More theorems on Hessenberg's sum}

    We need to prove some properties of $\oplus$, particularly about 
    its relation with the order $<$ on \texttt{T1}.

@@

    Associativity of Hessenberg's sum is proved the same way.

@@

    \paragraph*{\gaiasign} The module
      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
      compatible with \gaia's type \texttt{T1}.

    \subsection{A termination measure for hydra battles }

@@

    \paragraph*{\gaiasign}
    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.

@@

\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gcanon}).

\paragraph*{Remark}
In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.

@@

\paragraph*{\gaiasign} This lemma is also available in
Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:

\inputsnippets{GCanon/gcanonLt}

@@

\input{movies/snippets/Canon/canonSLimitLub}

@@

\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
\label{sect:pathes-intro}

@@

\paragraph*{\gaiasign}
The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gpath}).

@@

\section{A  proof of impossibility}

@@

\section{The case of standard battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

@@

\section{Length of minimal large sequences}

@@

\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.

>>>>

 \subsection*{Compatibility with \gaia (in progress)}
 \label{sect:gaia-first-intro}
The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
aims  to formalize mathematics in \coq  in
\index{gaiabridge}{Introduction}

the style of Nicolas Bourbaki. It contains many definitions and results about ordinal numbers. In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa. Remarks about compatibility with \gaia are signalled with the picture \gaiasign. A special index is in construction (page~\pageref{gaia-index}).

@@

\begin{itemize}
\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}. This representation uses several (hopefully harmless) axioms. We use it as a reference for proving the correctness of ordinal notations.

  \index{gaiabridge}{Introduction}
  \gaiasign
  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}. Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.

@@

@@

\subsection{About Gaia}
\index{gaiabridge}{Introduction}
\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.

@@

\input{movies/snippets/T1/finiteOrds}
\index{gaiabridge}{Finite ordinals}
\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.

@@

\index{gaiabridge}{The ordinal $\omega$}
\paragraph*{\gaiasign}
In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Exponential of base $\omega$}
In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Pretty printing Cantor normal forms}
In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Strict order on ordinals below $\epsilon_0$}

In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Ordinal terms in Cantor normal form}
In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Type of well formed ordinal terms below $\epsilon_0$}
 The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.

@@

\paragraph*{\gaiasign}
\index{gaiabridge}{Limit and successor ordinals}
In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.

@@

\subsection{An ordinal notation for  \gaia's ordinals}
\index{gaiabridge}{Ordinal notations}

\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).

\inputsnippets{T1Bridge/gEpsilon0}

@@

    Natural sum (Hessenberg  sum) is a commutative and monotonous version of
    addition. It is used as an auxiliary operation  for defining variants
    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.

@@

    \index{hydras}{Projects}
    \begin{project}
    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
    (\texttt{coq-equations}, \texttt{Function}, etc.).
    \end{project}
    \subsection{More theorems on Hessenberg sum}
\index{gaiabridge}{Hessenberg sum}
    We need to prove some properties of $\oplus$, particularly about 
    its relation with the order $<$ on \texttt{T1}.

@@

    Associativity of Hessenberg sum is proved the same way.

@@

    \paragraph*{\gaiasign} The module
      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg sum
      compatible with \gaia's type \texttt{T1}.
      \index{gaiabridge}{Hessenberg sum}
    \subsection{A termination measure for hydra battles }

@@

    \paragraph*{\gaiasign}
    \index{gaiabridge}{Termination of all hydra battles}
    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.

@@

\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gcanon}).
\index{gaiabridge}{Canonical sequences}
\paragraph*{Remark}
In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.

@@

\paragraph*{\gaiasign} This lemma is also available in
Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
\index{gaiabridge}{Canonical sequences}
\inputsnippets{GCanon/gcanonLt}

@@

\input{movies/snippets/Canon/canonSLimitLub}
\index{gaiabridge}{Canonical sequences}

@@

\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
\index{gaiabridge}{Accessibility and paths inside $\epsilon_0$}
\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
\label{sect:pathes-intro}

@@

\paragraph*{\gaiasign}

The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gpath}).

@@

\section{A  proof of impossibility}
\label{sect:impossibility-free}

@@

\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
\index{gaiabridge}{Impossibility theorems}
\section{The case of standard battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
\index{gaiabridge}{Impossibility theorems}

@@

\section{Length of minimal large sequences}
\label{sect:lalpha-section}

@@

\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
\index{gaiabridge}{Rapidly growing functions}
